Diversity and evolution of centromere repeats in the maize genome

### 2014-05-12 centc.tex
Push b4 submit, noted one error in supplementary stuf I could not fix, but still going in.

\author{Paul Bilinski \and Kevin Distor \and Jose Gutierrez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and Kelly Dawe \and Jeffrey Ross-Ibarra}

@@

\institute{Paul Bilinski : Kevin Distor : Jose Gutierrez-Lopez : Gabriela Mendoza Mendoza : Jeffrey Ross-Ibarra\at
              Department of Plant Sciences, UC Davis\\
              Davis, California, 95616
              \and
              Jeffrey Ross-Ibarra\at
              The Genome Center and Center for Population Biology, University of California\\
              Davis, California, 95616\\
              \email{rossibarra@ucdavis.edu}           %  \\
              \and
              Jose Gutierrez-Lopez\at
              Department of Natural Resources and the Environment
              University of New Hampshire\\
              Durham, New Hampshire, 03820
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Jinghua Shi : R. Kelly Dawe\at
           Department of Plant Biology, University of Georgia\\
	  Athens, Georgia, 30602
           \and
           R. Kelly Dawe\at
           Department of Genetics, University of Georgia\\
           Athens, Georgia, 30602\\
}

@@

Fluorescent in situ hybridization (FISH) was used to compare abundance of CentC repeats in chromosomes of a maize x \emph{parviglumis} F1 hybrid and a maize x \emph{Z. luxurians} F1 hybrid. \jri{Kellyto fill in}

@@

\begin{figure}
% For two-column wide figures use
% Use the relevant command to insert your figure file.
%  \includegraphics[width=0.75\textwidth]{circos.png}
\includegraphics{Fig1_circos}
\caption{CentC repeat location in relation to the maize subgenomes.  The outer ring depicts chromosomal assignment to the two subgenomes, with higher confidence regions are colored with darker colors.  Breakpoints between the subgenomes remain uncolored to indicate uncertainty. The middle ring, shaded in blue, displays the locations of all CentCs across the genome.  The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20KB in length}
\label{circos}    
\end{figure}

Previous studies have also described the Cent4 repeat, a tandem pericentromeric repeat that occurs only primarily \jri{correct? it does blast elsewhere?} on chromosome 4 \citep{Page2001}. 
Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002, jin2004maize}. 
BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.  
Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.  
But while previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}, RLX\_sela also appears to be missing the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a TE-derived tandem repeat unique to the pericentromere of chromosome 4.

@@

\begin{figure}
% For two-column wide figures use
% Use the relevant command to insert your figure file.
\centering
\includegraphics[width=1\textwidth]{Figure4_noedge}
\caption{(a) Mb of CentC in genomic libraries of maize and teosinte.  Box plots show data from \citet{Chia2012}.  Points show data for maize inbred B73 and the teosinte \emph{Z. luxurians} from \citet{Tenaillon2011}.  For comparison, the data point of maize inbred B73 in \citet{Chia2012} is shown with a tick mark on the box plot. (b) FISH images from an F1 hybrids between maize and the teosinte \emph{parviglumis}, and maize and the teosinte \emph{Z. luxurians}.  Upper row shows the karyotype for the teosinte and the lower row shows the maize karyotype.  Red=CentC, Green=CRM. 
}
\label{abundance}    
\end{figure}

@@

\jri{quick scan of the references looks OK, but do double check format etc.}

>>>>

\author{Paul Bilinski \and Kevin Distor \and Jose Gutierrez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and R. Kelly Dawe \and Jeffrey Ross-Ibarra}

@@

\institute{Paul Bilinski : Kevin Distor : Jose Gutierrez-Lopez : Gabriela Mendoza Mendoza : Jeffrey Ross-Ibarra\at
              Department of Plant Sciences, UC Davis\\
              Davis, California, 95616
              \and
              Jeffrey Ross-Ibarra\at
              The Genome Center and Center for Population Biology, University of California\\
              Davis, California, 95616\\
              \email{rossibarra@ucdavis.edu}           %  \\
              \and
              Jose Gutierrez-Lopez\at
              Department of Natural Resources and the Environment
              University of New Hampshire\\
              Durham, New Hampshire, 03820
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Jinghua Shi : R. Kelly Dawe\at
           Department of Plant Biology, University of Georgia\\
	  Athens, Georgia, 30602
}

@@

Fluorescent in situ hybridization (FISH) was carried out as described in \citet{kato2004paint} (\emph{Z. luxurians} hybrid) and \citet{Shi2010} (\emph{Z. parviglumis} hybrid).

@@

\begin{figure}
% For two-column wide figures use
% Use the relevant command to insert your figure file.
%  \includegraphics[width=0.75\textwidth]{circos.png}
\includegraphics{Fig1_circos}
\caption{CentC repeat location in relation to the maize subgenomes.  The outer ring depicts chromosomal assignment to the two subgenomes, with higher confidence regions are colored with darker colors.  Green corresponds to subgenome 1, brown to subgenome 2.  Breakpoints between the subgenomes remain uncolored to indicate uncertainty. The middle ring, shaded in blue, displays the locations of all CentCs across the genome.  The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20KB in length}
\label{circos}    
\end{figure}

Previous studies have also described the Cent4 repeat, a tandem pericentromeric repeat that occurs primarily on chromosome 4 \citep{Page2001}. 
Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002, jin2004maize}. 
BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.  
Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.  
But while previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}, RLX\_sela also appears to be missing the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a TE-derived tandem repeat unique to the pericentromere of chromosome 4.

@@

\begin{figure}
% For two-column wide figures use
% Use the relevant command to insert your figure file.
\centering
\includegraphics[width=1\textwidth]{Fig4_kelly2}
\caption{(a) FISH analysis of a single individual heterozygous for B73 and \emph{Zea luxurians} (GRIN accession PI422162).
Chromosomes, ordered from 1 (left) to 10 (right), were hybridized with the Birchler probe cocktail \citep{kato2004paint}. 
Green shows CentC and a 4-12-1 subtelomere repeat, blue the 180bp knob repeat, red the abundant TAG microsatellite and another subtelomeric repeat, white the TR1 knob repeat, orange the Cent4 repeat, yellow the 5S rDNA repeat, and aqua shows the Nucleolus Organizer Region.
The green signals at the primary constrictions (arrows) are CentC.
Note that \emph{Z. luxurians} has far brighter CentC signals.
This image was graciously provided by Patrice Albert.
(b) FISH analysis of a single individual heterozygous for B73 and \emph{Z. parviglumis}(GRIN accession PI566687).
Chromosomes were hybridized with the \citet{Shi2010} probe cocktail showing CentC in red and CRM2 in green. 
Each separate probe is shown separately below the two-color image, highlighting that CentC is more abundant in \emph{Z. parviglumis} and CRM2 is more abundant in maize. 
(c) Mb of CentC in genomic libraries of maize and teosinte. 
Box plots show data from \citet{Chia2012}. 
Points show data for maize inbred B73 and the teosinte \emph{Z. luxurians} from \citet{Tenaillon2011}. For comparison, the data point of maize inbred B73 in \citet{Chia2012} is shown with a tick mark on the box plot.
}
\label{abundance}    
\end{figure}

@@
